Bumble bee foraging bias

### 2021-01-14 code/visitation_bias.Rmd
update desktop

## Load data
```{r message=FALSE, warning=FALSE, include=FALSE, echo = FALSE}
# Site data
site_data <- read_csv("../processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
  mutate(elev.class2 = factor(elev.class2, levels = c("low", "mid", "high")),
         site = factor(site)) # turn elav.class into an ordered factor and site into a regular factor

# BB traits
bb_traits <- read_csv("../processed_data/bb_traits.csv")

# FL traits
## family-level taxonomy
fl_tax <- read_csv("../processed_data/floral_tax.csv")

## Kugler morphotypes (accessed through Bioflor)
fl_ktype <- read_csv("../processed_data/floral_k_type.csv")

## Flower colors (accessed through BioFlor)
fl_color <- read_csv("../processed_data/floral_color.csv")

## diameter and corolla depth data from Gita Benadi
fl_morph <- read_csv("../processed_data/flower_morphology_GitaBenadi.csv") %>% 
  dplyr::select(SampleID, plant.family = Family, plant.sp = Species, 
                Unit, Flowers_per_unit, diam = Diameter, LN) %>%
  group_by(plant.sp) %>%
  summarize(diam = mean(diam), LN = mean(LN))

## verbal description of Kugler morphotypes
fl_ktype_key <- read_delim("../processed_data/kugler_key.tsv", delim = "\t", col_names = FALSE) %>%
  dplyr::select(k.type = 1, description = 2)

@@

## Align survey and visitation data with pseudodates
```{r}
# Site-dates where BB visitation was recorded (excluding dates where only males were recorded)
visitation_samples <- read_csv("../processed_data/network.csv") %>%
  filter(caste != "male") %>%
  select(site, date) %>%
  distinct()

# Site dates where floral surveying was conducted
survey_samples <- read_csv("../processed_data/floral_survey.csv") %>%
  select(site, date) %>%
  distinct()

@@

## Generate a data frame that represents visitation and survey abundance
```{r}
# BB visitation data
visitation <- read_csv("../processed_data/network.csv") %>%
  filter(caste != "male") %>% # we will ignore males 
  group_by(site, date, bb.sp, plant.sp) %>%
  summarize(visits = n()) %>% # visits per BB-FL pair
  ungroup() %>%
  left_join(pseudodates) %>% # Now we will replace unmatched dates with their pseudo dates
  replace_na(replace = list(mismatch = 0)) %>% # set mismatch to zero for perfectly aligned dates
  filter(mismatch <= 7) %>% # this gets rid of irreconcilable dates (like the t4 dates mentioned above)
  mutate(date2 = if_else(is.na(pseudo.date), date, pseudo.date)) %>% # replace date with pseudodate for all site-dates lacking perfect alignments
  select(-c(date, pseudo.date, mismatch)) %>%
  rename(date = date2) %>%
  group_by(site, date) %>%
  complete(bb.sp, plant.sp, fill = list(visits = 0)) # generate all possible BB-FL combos for each site date and fill in zeros for visits

# Floral cover data
flower_cover <- read_csv("../processed_data/floral_survey.csv") %>%
  filter(flower.cover > 0) %>% # ignore plants that were not blooming
  group_by(site, date, plant.sp) %>%
  summarize(flower.cover = sum(flower.cover))

@@

# 1. How does overall bias (Blüthgen's d') vary across species and through elev*time? For now, we will focus on bias at the level of plant species. We'll start considering traits later on when our goal is to explain patterns of bias.
## 1.1 Calculate Blüthgen's d' (i.e. visitation bias)
```{r}
# No grouping -- just plant.sp
bias_plantsp <- biaseval(visit_cover, plant.sp, site, date) %>%
  left_join(site_data) %>% # add site data
  left_join(bb_abundance) %>% # add BB abundance data
  left_join(fl_abundance) %>% # add fl abundance
  mutate(criterion = factor(rep("plant.sp", n())))

@@

# This model includes a global interaction term for yday and elevation, with a varying intercept for bb.sp and a varying slow for elev | bb.sp. I think this should give me all the estimation I'm interested in.
brm.plantsp11 <- brm(dprime ~ yday * elev + (1 + elev | bb.sp), 
                  control = list(max_treedepth = 10),
                  cores = 4,
                  iter = 4000,
                  #family = "skew_normal",
                  file = "./output/brm_plantsp11",
                  data = filter(bias, criterion ==  "plant.sp"))

@@

brm.plantsp1 <- brm(dprime ~ yday * elev + (1 | bb.sp), 
                  control = list(max_treedepth = 10),
                  cores = 4,
                  iter = 2000,
                  file = "./output/brm_plantsp1",
                  data = filter(bias, criterion ==  "plant.sp"))

@@

brm.plantsp2 <- brm(dprime ~ yday + elev + (1 | bb.sp), 
                  control = list(max_treedepth = 10),
                  cores = 4,
                  iter = 2000,
                  file = "./output/brm_plantsp2",
                  data = filter(bias, criterion ==  "plant.sp"))

@@

# This, I think, would also be a legitimate way to specify the model, but it would not tell us anything about elevation or time. 
brm.plantsp0 <- brm(dprime ~ 1 + (1 | bb.sp) + (1 | site),
                  control = list(max_treedepth = 10),
                  cores = 4,
                  iter = 2000,
                  file = "./output/brm_plantsp0",
                  data = filter(bias, criterion ==  "plant.sp"))

summary(brm.plantsp0)
plot(brm.plantsp0)
parnames(brm.plantsp0)
mcmc_plot(brm.plantsp0, pars = c("^r_"))
mcmc_plot(brm.plantsp0, pars = c("^b_"))
pp_check(brm.plantsp0)
bayes_R2(brm.plantsp0)
```

@@

# 2. Having established the bias of BB foraging, let's zoom in on which morphotypes and families were over/under-represented

@@

@@

## 2.3 Calculate bias per plant speces/ktype/family
```{r}
representation <- biaseval_group(visit_cover, plant.sp) %>%
  rename(plant.sp = criterion) %>%
  left_join(site_data) %>%
  left_join(fl_traits) %>%
  mutate(k.type.ss = factor(k.type.ss)) %>%
  select(site, date, elev.mean, yday, bb.sp, plant.sp, k.type.ss, 
         prop.visits, prop.flower.cover, bias, visits, 
         flower.cover, plant.genus, plant.family)

# 
representation_ktype <- biaseval_group(visit_cover, k.type.ss) %>%
  left_join(site_data) %>%
  rename(group = criterion) %>%
  mutate(criterion = rep("ktype", n()))
# 
# bias_group_family <- biaseval_group(visit_cover, plant.family) %>%
#   left_join(site_data) %>%
#   rename(group = criterion) %>%
#   mutate(criterion = rep("plant.family", n()))
# 
# bias_group <- bind_rows(bias_group_plantsp, bias_group_ktype, bias_group_family)
# 
# ggplot(bias_group, aes(prop.flower.cover, prop.visits)) +
#   geom_point(alpha = 0.1) +
#   #geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
#   geom_smooth(method = "lm") +
#   geom_abline(slope = 1, color = "red", linetype = "dashed") +
#   facet_wrap(~criterion)
# 
# ggplot(bias_group, aes(log(prop.flower.cover), prop.visits)) +
#   geom_point(alpha = 0.1) +
#   geom_smooth() +
#   facet_wrap(~criterion)
# 
# ggplot(filter(bias_group, criterion == "ktype"), aes(prop.flower.cover, prop.visits)) +
#   geom_point(alpha = 0.1) +
#   geom_smooth() +
#   geom_abline(slope = 1, color = "red", linetype = "dashed") +
#   facet_wrap(~group)
```

@@

## 2.5 Explaining bias by floral traits, BB spp, and elevation (and maybe time)
```{r}
ggplot(representation, aes(k.type.ss, bias)) +
  geom_violin()

ggplot(representation_ktype, aes(group, bias)) +
  geom_boxplot()

@@

ggplot(filter(representation, visits > 0), aes(bias)) +
  geom_histogram() +
  facet_wrap(~k.type.ss)

@@

summary(lm(representation ~ k.type.ss, data = bias))

brm_representation0 <- brm(bias ~ 1 + (1 | group),
                 control = list(max_treedepth = 10),
                 cores = 4,
                 iter = 2000,
                 file = "./output/brm_representation1",
                 data = representation_ktype)

brm_representation1 <- brm(bias ~ 1 + (1 | k.type.ss),
                 control = list(max_treedepth = 10),
                 #cores = 4,
                 cores = 4,
                 chains = 4,
                 #iter = 2000,
                 iter = 2000,
                 file = "./output/brm_representation1",
                 data = representation)

@@

summary(lmer_representation3)
plot(lmer_representation3)
```

# 3. How do species vary in their biases?
```{r}
rep_matrix.plantsp <- representation %>%
  group_by(bb.sp, group) %>%
  summarize(bias = mean(bias)) %>%
  pivot_wider(values_from = bias, names_from = group, values_fill = 0) %>%
  column_to_rownames("bb.sp")

bias_matrix.ktype <- bias_group_ktype %>%
  group_by(bb.sp, group) %>%
  summarize(bias = mean(bias)) %>%
  pivot_wider(values_from = bias, names_from = group, values_fill = 0) %>%
  column_to_rownames("bb.sp")

@@

pca.plantsp <- rda(bias_matrix.plantsp)
plot(pca.plantsp, type = "t")
envfit(pca.plantsp ~ pbl.w.class, data = traits)
summary(ordisurf(pca.plantsp ~ pbl.w.class, data = traits))

pca.ktype <- rda(bias_matrix.ktype)
plot(pca.ktype, type = "t")
envfit(pca.ktype ~ pbl.w.class, data = traits)
summary(ordisurf(pca.ktype ~ pbl.w.class, data = traits))
pca.ktype_pca1_traits <- pca.ktype$CA$u[,1] %>%
  as.data.frame() %>%
  rownames_to_column(var = "bb.sp") %>%
  rename(PC1 = 2) %>%
  left_join(bb_traits)

ggplot(pca.ktype_pca1_traits, aes(pbl.w, PC1)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ poly(x, 2))

ggplot(pca.ktype_pca1_traits, aes(pbl.w.class, PC1)) +
  geom_point() +
  geom_smooth(method = "lm")
```

# 3. Can we explain bias by BB-Fl trait associations? (RLQ + fourth corner analysis)
## 3.1
## 3.2 Do BB-FL trait associations vary with elevation?

# 4. How does bias relate to network structure?
## 4.1 A function for calculating niche overlap
```{r}
# Niche overlap by sp.pair * site * year
niche_overlapper <- function(x, criterion) {

@@

## 4.2 Calculate niche overlap: NOT READY FOR INTERPRETATION YET
```{r}
niche_overlap_plant.sp <- niche_overlapper(visit_cover, plant.sp)

@@

>>>>

## Load data
```{r message=FALSE, warning=FALSE, include=FALSE, echo = FALSE}
# Site data
site_data <- read_csv("../data/processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
  mutate(elev.class2 = factor(elev.class2, levels = c("low", "mid", "high")),
         site = factor(site)) # turn elav.class into an ordered factor and site into a regular factor

# BB traits
bb_traits <- read_csv("../data/processed_data/bb_traits.csv")

# FL traits
## family-level taxonomy
fl_tax <- read_csv("../data/processed_data/floral_tax.csv")

## Kugler morphotypes (accessed through Bioflor)
fl_ktype <- read_csv("../data/processed_data/floral_k_type.csv")

## Flower colors (accessed through BioFlor)
fl_color <- read_csv("../data/processed_data/floral_color.csv")

## diameter and corolla depth data from Gita Benadi
fl_morph <- read_csv("../data/processed_data/flower_morphology_GitaBenadi.csv") %>% 
  dplyr::select(SampleID, plant.family = Family, plant.sp = Species, 
                Unit, Flowers_per_unit, diam = Diameter, LN) %>%
  group_by(plant.sp) %>%
  summarize(diam = mean(diam), LN = mean(LN))

## verbal description of Kugler morphotypes
fl_ktype_key <- read_delim("../data/processed_data/kugler_key.tsv", delim = "\t", col_names = FALSE) %>%
  dplyr::select(k.type = 1, description = 2)

@@

## Align survey and visitation data with pseudodates
```{r}
# Site-dates where BB visitation was recorded (excluding dates where only males were recorded)
visitation_samples <- read_csv("../data/processed_data/network.csv") %>%
  filter(caste != "male") %>%
  select(site, date) %>%
  distinct()

# Site dates where floral surveying was conducted
survey_samples <- read_csv("../data/processed_data/floral_survey.csv") %>%
  select(site, date) %>%
  distinct()

@@

## Generate a data frame that represents visitation and survey abundance
```{r}
# BB visitation data
visitation <- read_csv("../data/processed_data/network.csv") %>%
  filter(caste != "male") %>% # we will ignore males 
  group_by(site, date, bb.sp, plant.sp) %>%
  summarize(visits = n()) %>% # visits per BB-FL pair
  ungroup() %>%
  left_join(pseudodates) %>% # Now we will replace unmatched dates with their pseudo dates
  replace_na(replace = list(mismatch = 0)) %>% # set mismatch to zero for perfectly aligned dates
  filter(mismatch <= 7) %>% # this gets rid of irreconcilable dates (like the t4 dates mentioned above)
  mutate(date2 = if_else(is.na(pseudo.date), date, pseudo.date)) %>% # replace date with pseudodate for all site-dates lacking perfect alignments
  select(-c(date, pseudo.date, mismatch)) %>%
  rename(date = date2) %>%
  group_by(site, date) %>%
  complete(bb.sp, plant.sp, fill = list(visits = 0)) # generate all possible BB-FL combos for each site date and fill in zeros for visits

# Floral cover data
flower_cover <- read_csv("../data/processed_data/floral_survey.csv") %>%
  filter(flower.cover > 0) %>% # ignore plants that were not blooming
  group_by(site, date, plant.sp) %>%
  summarize(flower.cover = sum(flower.cover))

@@

# 1. Quantifying bias
How does overall bias (Blüthgen's d') vary across species and through elev*time? For now, we will focus on bias at the level of plant species. We'll start considering traits later on when our goal is to explain patterns of bias.
## 1.1 Calculate Blüthgen's d' (i.e. visitation bias)
```{r}
# No grouping -- just plant.sp
bias_plantsp <- biaseval(visit_cover, plant.sp, site, date) %>%
  left_join(site_data) %>% # add site data
  left_join(bb_abundance) %>% # add BB abundance data
  left_join(fl_abundance) %>% # add fl abundance
  mutate(criterion = factor(rep("plant.sp", n())))

@@

# This model includes a global interaction term for yday and elevation, with a varying intercept for bb.sp and a varying slow for elev | bb.sp. I think this should give me all the estimation I'm interested in. It's tricky to make it converge, though.
brm.plantsp11 <- brm(dprime ~ yday * elev + (1 + elev | bb.sp), 
                  control = list(max_treedepth = 10),
                  cores = 4,
                  iter = 4000,
                  #family = "skew_normal",
                  file = "../output/brm_plantsp11",
                  data = filter(bias, criterion ==  "plant.sp"))

@@

# Alternatively, we can drop the varying slope term to simplify the model a bit. Still doesn't quite converge.
brm.plantsp1 <- brm(dprime ~ yday * elev + (1 | bb.sp), 
                  control = list(max_treedepth = 10),
                  cores = 4,
                  iter = 2000,
                  file = "../output/brm_plantsp1",
                  data = filter(bias, criterion ==  "plant.sp"))

@@

# We can simplify the model still more by dropping the yday*elev interaction term. Now it converges. But I think the takeaway is that neither elevation noy yday give us a clear signal with respect  to dprime.
brm.plantsp2 <- brm(dprime ~ yday + elev + (1 | bb.sp), 
                  control = list(max_treedepth = 10),
                  cores = 4,
                  iter = 2000,
                  file = "../output/brm_plantsp2",
                  data = filter(bias, criterion ==  "plant.sp"))

@@

# This, I think, would also be a legitimate way to specify the model, but it would not tell us anything about elevation or time. It's the simplest, and that's nice. Well-behaved. Captures about twice as much variance as the previous model, and it wins the WAIC context easily.
brm.plantsp0 <- brm(dprime ~ 1 + (1 | bb.sp) + (1 | site),
                  control = list(max_treedepth = 10),
                  cores = 4,
                  iter = 2000,
                  file = "../output/brm_plantsp0",
                  data = filter(bias, criterion ==  "plant.sp"))

summary(brm.plantsp0)
plot(brm.plantsp0)
parnames(brm.plantsp0)

# I guess the message here is that both bb.sp and site contribute to the variation of dprime, but site more so. Biologically, I think the meaning of prime variation by site is that some sites have greater relative abundance of *attractive* flora than others. A site with loads of something that is never visited and a little bit of clover will yield a high dprime, while a site with a high abundance of attractive floral would yield a low dprime.
mcmc_plot(brm.plantsp0, pars = c("^r_"))
mcmc_plot(brm.plantsp0, pars = c("^r_bb.sp")) 
ggsave("../output/brm.plantsp0_plot1.png")
mcmc_plot(brm.plantsp0, pars = c("^b_"))
pp_check(brm.plantsp0)
bayes_R2(brm.plantsp0)

# Another option would be to drop just elevation from the model. This runs clean. Doesn't explain much more variance than the simpler model, though.
brm.plantsp00 <- brm(dprime ~ yday + (1 | bb.sp) + (1 | site),
                  control = list(max_treedepth = 10),
                  cores = 4,
                  iter = 2000,
                  file = "../output/brm_plantsp00",
                  data = filter(bias, criterion ==  "plant.sp"))

summary(brm.plantsp00)
plot(brm.plantsp00)
parnames(brm.plantsp00)
mcmc_plot(brm.plantsp00, pars = c("^r_"))
mcmc_plot(brm.plantsp00, pars = c("^b_"))
conditional_effects(brm.plantsp00)
pp_check(brm.plantsp00)
bayes_R2(brm.plantsp00)

# Let's do a formal comparison of the last three models
waic(brm.plantsp2, brm.plantsp0, brm.plantsp00)

# The one including yday is a *slight* winner, but the margin is so small as to be meaningless. I think the simplest model is the one we want (brm.plantsp0).

# Let's try the same model now, but for ktype.
brm.ktype0 <- brm(dprime ~ 1 + (1 | bb.sp) + (1 | site),
                  control = list(max_treedepth = 10),
                  cores = 4,
                  iter = 2000,
                  family = "skew_normal",
                  file = "../output/brm_ktype0",
                  data = filter(bias, criterion ==  "ktype"))

summary(brm.ktype0)
plot(brm.ktype0)
parnames(brm.ktype0)
mcmc_plot(brm.ktype0, pars = c("^r_"))
mcmc_plot(brm.ktype0, pars = c("^b_"))
conditional_effects(brm.ktype0)
pp_check(brm.ktype0)
bayes_R2(brm.ktype0)
```

@@

# 2. Explaining bias

@@

*We will adopt the term representation to avoid confusion with the object defined above.*

@@

## 2.3 Calculate bias per plant speces/ktype/family
```{r}
representation <- biaseval_group(visit_cover, plant.sp) %>%
  rename(plant.sp = criterion) %>%
  left_join(site_data) %>%
  left_join(fl_traits) %>%
  mutate(k.type.ss = factor(k.type.ss)) %>%
  select(site, date, elev.mean, yday, bb.sp, plant.sp, k.type.ss, 
         prop.visits, prop.flower.cover, bias, visits, 
         flower.cover, plant.genus, plant.family) %>%
  left_join(bb_traits)

# 
# representation_ktype <- biaseval_group(visit_cover, k.type.ss) %>%
#   left_join(site_data) %>%
#   rename(group = criterion) %>%
#   mutate(criterion = rep("ktype", n()))
# 
# bias_group_family <- biaseval_group(visit_cover, plant.family) %>%
#   left_join(site_data) %>%
#   rename(group = criterion) %>%
#   mutate(criterion = rep("plant.family", n()))
# 
# bias_group <- bind_rows(bias_group_plantsp, bias_group_ktype, bias_group_family)
# 
# ggplot(bias_group, aes(prop.flower.cover, prop.visits)) +
#   geom_point(alpha = 0.1) +
#   #geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
#   geom_smooth(method = "lm") +
#   geom_abline(slope = 1, color = "red", linetype = "dashed") +
#   facet_wrap(~criterion)
# 
# ggplot(bias_group, aes(log(prop.flower.cover), prop.visits)) +
#   geom_point(alpha = 0.1) +
#   geom_smooth() +
#   facet_wrap(~criterion)
# 
# ggplot(filter(bias_group, criterion == "ktype"), aes(prop.flower.cover, prop.visits)) +
#   geom_point(alpha = 0.1) +
#   geom_smooth() +
#   geom_abline(slope = 1, color = "red", linetype = "dashed") +
#   facet_wrap(~group)
```

@@

## 2.5 Multi-level models
```{r}
# The tons of zeros are due to the lots of plant species that were present at very low abundance and received few or zero visits. 
ggplot(representation, aes(k.type.ss, bias)) +
  geom_violin()

ggplot(representation, aes(k.type.ss, bias)) +
  geom_boxplot()

@@

ggplot(filter(representation, abs(bias) > 0.1), aes(bias)) +
  geom_histogram() +
  facet_wrap(~k.type.ss)

@@

brm_representation0 <- brm(bias ~ 1 + (1 | k.type.ss) + (1 | site),
                 control = list(max_treedepth = 10),
                 cores = 4,
                 chains = 4,
                 iter = 2000,
                 file = "../output/brm_representation0",
                 data = representation)

summary(brm_representation0)
plot(brm_representation0)
parnames(brm_representation0)
mcmc_plot(brm_representation0, pars = c("^r_"))
#mcmc_plot(representation.ktype0, pars = c("elev"))
mcmc_plot(brm_representation0, pars = c("^b_"))
mcmc_plot(brm_representation0)
#conditional_effects(representation.ktype0)
pp_check(brm_representation0)
bayes_R2(brm_representation0)

brm_representation1 <- brm(bias ~ pbl.w + (1 + pbl.w | k.type.ss) + (1 | bb.sp),
                 control = list(max_treedepth = 10),
                 cores = 4,
                 chains = 4,
                 iter = 2000,
                 file = "../output/brm_representation1",
                 data = representation)

@@

summary(lmer_representation3)
plot(lmer_representation3)

# Can we explain anything with tongue length

ggplot(filter(representation, bb.sp != "gers" & abs(bias) >= 0.1), aes(pbl.w, bias)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~k.type.ss)

lm_pbl <- lmer(bias ~ pbl.w + (1 + pbl.w | k.type.ss) + (1 | bb.sp),
               data = representation)

summary(lm_pbl)
```

## 2.6 Ordination of bias
```{r}
rep_matrix.ktype <- representation %>%
  group_by(bb.sp, k.type.ss) %>%
  summarize(bias = mean(bias)) %>%
  pivot_wider(values_from = bias, names_from = k.type.ss, values_fill = 0) %>%
  column_to_rownames("bb.sp")

@@

pca.ktype <- rda(rep_matrix.ktype)
#ef.ktype <- envfit(pca.ktype ~ pbl.w.class, data = traits)
#os.ktype <- ordisurf(pca.ktype ~ pbl.w.class, data = traits)
ef.ktype <- envfit(pca.ktype ~ pbl.w, data = traits)
os.ktype <- ordisurf(pca.ktype ~ pbl.w, data = traits)
plot(pca.ktype, type = "t")
plot(ef.ktype, add = TRUE)
plot(os.ktype)
```

# 3. Effects of bias

## 3.1 Relationship between bias(d') and niche overlap
### 3.1.1 A function for calculating niche overlap
```{r}
# Niche overlap by sp.pair * site * year
niche_overlapper <- function(x, criterion) {

@@

### 3.1.2 Calculate niche overlap: NOT READY FOR INTERPRETATION YET
```{r}
niche_overlap_plant.sp <- niche_overlapper(visit_cover, plant.sp)

@@

## 3.2 Ordination of dietary niche
```{r}
rep_matrix.ktype <- representation %>%
  group_by(bb.sp, k.type.ss) %>%
  summarize(bias = mean(bias)) %>%
  pivot_wider(values_from = bias, names_from = k.type.ss, values_fill = 0) %>%
  column_to_rownames("bb.sp")

pca.ktype <- rda(rep_matrix.ktype)
#ef.ktype <- envfit(pca.ktype ~ pbl.w.class, data = traits)
#os.ktype <- ordisurf(pca.ktype ~ pbl.w.class, data = traits)
ef.ktype <- envfit(pca.ktype ~ pbl.w, data = traits)
os.ktype <- ordisurf(pca.ktype ~ pbl.w, data = traits)
plot(pca.ktype, type = "t")
plot(ef.ktype, add = TRUE)
plot(os.ktype)
```

### 2021-01-14 writing/manuscript_v4.Rmd
update desktop

Bumble bees (Hymenoptera: *Bombus* spp.) are an especially interesting group in which to study floral biases together with their causal mechanisms and downstream effects. On the one hand, bumble bees are exceptionally versatile foragers, and their ability to solve the puzzles posed by complex flower morphologies [@Plowright1984-lc], or indeed by human investigators [@Alem2016-rn; @Loukola2017-bt], renders almost any flower potentially accessible. Even when a gross mismatch between tongue length and corolla depth makes the nectar of a given flower inaccessible by ordinary means, some bumble bee species will resort readily to biting through the base of the flower to “rob” nectar, incidentally making the nectar available to other prospective foragers as well. Flower handling skills are, moreover, socially transmitted, amplifying the plasticity of individual behavior [@Leadbeater2008-rh; @Loukola2017-bt]. Nevertheless, the flexibility of bumble bee foraging is tuned to floral reward, which depends on both the intrinsic reward of a given flower and the searching/learning/handling costs associated with exploiting it. At an evolutionary level, the innate preferences of bumble bees for certain floral traits likely reflect historic associations between floral traits and floral rewards [@Raine2007-nq]. These innate preferences, however, can be modulated or even nullified by patterns of reward learned during foraging [@Kugler1943-uy; Maharaj2019-wm], which depend not only intrinsic floral properties but also ecological modulators, such as the spatial distribution of floral patches or competition with other bumble bees [@Ishii2013-sm; @Plowright1984-lc; @Sowig1989-ld]. Thus, realized biases in bumble bee foraging are the labile product of bumble bee traits, floral traits, and ecological context. The influence of these biases on interaction network structure has consequences both for plant pollination and bumble bee species coexistence, and these consequences potentially feed back into the causes of bumble bee foraging bias via the modulation of competition between bumble bees and ultimately via the regulation of community assembly.  <!-- strengthen this sentence with some refs? -->

The topic of floral preference in bumble bees has been approached via both laboratory experimentation [e.g. @Kugler1943-uy; @Raine2007-nq] and observational field study [e.g. Heinrich1976-wc; @Goulson2008-zy]. The former approach enables more confident causal inference with respect to the mechanisms of floral choice, but it suffers all the usual limitations of biological realism associated with experimental control, and the higher-order effects of floral choices on network structure cannot be observed meaningfully under laboratory conditions. Observational field studies, however, have tended to be anecdotal in scope, typically describing bumble bee foraging patterns in particular localities, over short time frames, and with limited floristic data to enable a distinction between true foraging bias and the neutral effects of floral abundance [though see @Neumayer1999-il]. The steep elevation gradients of mountain slopes provide an opportunity for more powerful observational study of bumble foraging because they generate broad gradients of floristic and abiotic variation while controlling for confounding effects of geographic variation, such as genotypic differentiation within bumble bee species. When bumble bee visitation in such environments is sampled with sufficient spatial and temporal replication and complemented with quantitative floral surveying, inference can be made regarding both the trait interactions that underpin floral biases and the network structures that result from them. Mountain habitats are also arguably the most relevant places to study bumble bee ecology, since bumble bees are adapted to cool climates, and their global distribution is concentrated in areas of high elevation or high latitude. In such environments, bumble bees tend to dominate the pollinator community, and local flora often depend primarily on bumble bee visitation for pollination <!-- This paragraph needs to be beefed up a bit, but it at least sketches out the points I want to make -->.

@@

# Results

>>>>

Bumble bees (Hymenoptera: *Bombus* spp.) are an especially interesting group in which to study floral biases together with their causal mechanisms and downstream effects. On the one hand, bumble bees are exceptionally versatile foragers, and their ability to solve the puzzles posed by complex flower morphologies [@Plowright1984-lc], or indeed by human investigators [@Alem2016-rn; @Loukola2017-bt], renders almost any flower potentially accessible. Even when a gross mismatch between tongue length and corolla depth makes the nectar of a given flower inaccessible by ordinary means, some bumble bee species will resort readily to biting through the base of the flower to “rob” nectar, incidentally making the nectar available to other prospective foragers as well. Flower handling skills are, moreover, socially transmitted, amplifying the plasticity of individual behavior [@Leadbeater2008-rh; @Loukola2017-bt]. Nevertheless, the flexibility of bumble bee foraging is tuned to floral reward, which depends on both the intrinsic reward of a given flower and the searching/learning/handling costs associated with exploiting it. At an evolutionary level, the innate preferences of bumble bees for certain floral traits likely reflect historic associations between floral traits and floral rewards [@Raine2007-nq]. These innate preferences, however, can be modulated or even nullified by patterns of reward learned during foraging [@Kugler1943-uy; Maharaj2019-wm], which depend not only on intrinsic floral properties but also ecological modulators, such as the spatial distribution of floral patches or competition with other bumble bees [@Ishii2013-sm; @Plowright1984-lc; @Sowig1989-ld]. Thus, realized biases in bumble bee foraging are the labile product of bumble bee traits, floral traits, and ecological context. The influence of these biases on interaction network structure has consequences both for plant pollination and bumble bee species coexistence, and these consequences potentially feed back upon the causes of bumble bee foraging bias via the modulation of competition between bumble bees and ultimately via the regulation of community assembly.  <!-- strengthen this sentence with some refs? -->

The topic of floral preference in bumble bees has been approached via both laboratory experimentation [e.g. @Kugler1943-uy; @Raine2007-nq] and observational field study [e.g. Heinrich1976-wc; @Goulson2008-zy]. The former approach enables stronger causal inference with respect to the mechanisms of floral choice, but it suffers all the usual limitations of biological realism associated with experimental control, and the higher-order effects of floral choices on network structure cannot be observed meaningfully under laboratory conditions. Observational field studies, however, tend to be anecdotal in scope, typically describing bumble bee foraging patterns in particular localities, over short time frames, and with limited floristic data to enable a distinction between true foraging bias and the neutral effects of floral abundance [though see @Neumayer1999-il]. The steep elevation gradients of mountain slopes provide an opportunity for more powerful observational study of bumble foraging because they generate broad gradients of floristic and abiotic variation while controlling for confounding effects of geographic variation, such as genotypic differentiation within bumble bee species. When bumble bee visitation in such environments is sampled with sufficient spatial and temporal replication and complemented with quantitative floral surveying, inference can be made regarding both the trait interactions that underpin floral biases and the network structures that result from them. Mountain habitats are also arguably the most relevant places to study bumble bee ecology, since bumble bees are adapted to cool climates, and their global distribution is concentrated in areas of high elevation or high latitude. In such environments, bumble bees tend to dominate the pollinator community, and local flora often depend primarily on bumble bee visitation for pollination. <!-- This paragraph needs to be beefed up a bit, but it at least sketches out the points I want to make. I remember a reference somewhere -- I think in one of the Nordic papers, and I think a recent one -- that suggested that the morphology of alpine flowers has been shaped by selection pressure to attract bumble bees -->

@@

# Results
## Cumulative bias across species and sites
```{r}
knitr::include_graphics("../output/brm.plantsp0_plot1.png")
```

## Parsing bias by floral morphotype [representation ~ 1 + (1 + ktype) + (1 + site)]

## Linking morphological bias to species' traits [bias ordination w/ tongue length]

##
